---
title: "BIS15L_AI_class"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?
AI, or artificial intelligence, refers to the simulation of human intelligence processes by machines, particularly computer systems. This includes tasks such as learning, reasoning, problem-solving, perception, and language understanding. AI technology is used in a wide range of applications, including virtual assistants, self-driving cars, medical diagnosis, and predictive analytics.

2. What is ChatGPT?
ChatGPT is an AI-powered chatbot developed by OpenAI that is designed to have conversations with users in a more conversational and human-like manner. It uses a large language model to generate responses and engage in discussions on a wide range of topics. ChatGPT is designed to be used for entertainment, information, or assistance with tasks.

3. What is Github Copilot?
Github Copilot is an AI-powered code completion tool developed by Github in collaboration with OpenAI. It helps developers write code faster by providing suggestions and completing code snippets based on context and patterns found in the codebase. Copilot supports a wide range of programming languages and can generate entire functions or classes based on a description or comment provided by the developer. It aims to improve the efficiency and productivity of developers by automating repetitive coding tasks and reducing the need to search for documentation or examples.

4. Do you know the muffin man?
Yes, I know the muffin man! The muffin man lives on Drury Lane.

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix
netflix <- read_csv("data/netflix_titles.csv")


```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice

Use names to get the column names of the netflix data.
```{r}
names(netflix)
```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*


Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
netflix_movies <- netflix %>% 
  filter(type=="Movie")
```


Which director has the most movies on netflix?
```{r}
netflix_movies %>% 
  filter(!director=="NA") %>% 
  count(director, sort=T) %>% 
  arrange(desc(n))

```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

[insert your prompt here]
```{r}
netflix_movies %>% 
  separate_rows(director, sep = ",") %>% 
  filter(!director=="NA") %>% 
  count(director, sort=T) %>% 
  arrange(desc(n))
```



Which director had the most movies by release year and remove the NA's?
```{r}


```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


[Insert your prompt here]
```{r}

```



Make a bar plot of the top directors with the most movies on netflix
```{r}
netflix_movies %>% 
  ggplot(aes(x="director"))+
  geom_bar(position="dodge")
```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

[Insert your prompt here]
```{r}

```

*How about for just 2020*



[Insert your prompt here]
```{r}

```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt


*Now use ChatGPT, Copilot or previous notes to fix the code*





## That's it! Let's take a break and then move on to working on group projects.
